---
title: "Exploratory Data Analysis"
author: "Matthew Nguyen"
date: "20/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries and data

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)

df <- read.csv('../data/variants.csv')
```

## Distribution of frequencies

```{r}
ggplot(df, aes(x=frequency)) + 
  geom_histogram()
```

We see that most of the variant frequencies are very low. Let's look closer at the distribution there.

```{r}
ggplot(df[df[, 'frequency'] < 0.01,], aes(x=frequency)) + 
  geom_histogram()
```

Now let's look at the frequencies for each variant impact level

```{r}
ggplot(df[df[, 'impact'] == 'HIGH',], aes(x=frequency)) + 
  geom_histogram() + 
  labs(title='High impact')
```

```{r}
ggplot(df[df[, 'impact'] == 'MODERATE',], aes(x=frequency)) + 
  geom_histogram() + 
  labs(title='Moderate impact')
```

```{r}
ggplot(df[df[, 'impact'] == 'LOW',], aes(x=frequency)) + 
  geom_histogram() + 
  labs(title='Low impact')
```

```{r}
ggplot(df[df[, 'impact'] == 'MODIFIER',], aes(x=frequency)) + 
  geom_histogram() + 
  labs(title='Intergenic impact')
```

Now let's look at the distribution of estimated risk probability for each variant (calculated by the number of severe cases divided by the number of cases containing the variant, because of the binomial distribution). This will help us choose a prior.

```{r}
df %>%
  mutate(prob = num_severe / num_freq) %>%
  ggplot(aes(x=prob)) + 
    geom_histogram() + 
    labs(x='numberOfSevereCases/numberOfCasesWithVariant')
```
The distribution has many irregularities, with a large peak at 1. What if we do this for variants that are in at least 1% of samples?

```{r}
df[df[, 'frequency'] > 0.01,] %>%
  mutate(prob = num_severe / num_freq) %>%
  ggplot(aes(x=prob)) + 
    geom_histogram() + 
    labs(x='numberOfSevereCases/numberOfCasesWithVariant')
```
